|

### 2021-12-11 text/ManuscriptDraft.Rmd
Updated manuscript (introduction)

---
new_session: FALSE
title: |
  **Monitoring the phenology of individual flowers using deep learning and automatic tracking**
author:
- Hjalte M. R. Mann
- Alexandros Iosifidis
- Toke T. Høye
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
fontfamily: mathpazo
fontsize: 12pt
geometry: margin = 1in
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{dcolumn}
- \usepackage{caption}
- \usepackage{float}
- \usepackage{afterpage}
- \usepackage{siunitx}
- \usepackage{amsmath}
keywords: Citizen science, Arctic, insects
bibliography: ./library.bib
csl: ./journal-of-ecology.csl
abstract: ABSTRACT | Often simple variables will be used to describe the flowering phenology of a population of plants, e.g. onset or peak of flowering and to infer respones to climate change. Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of indiviuals. Further, we present an automatic flower tracking algorithm.
---

@@

What questions do we want to answer?

*Flower information*

@@

The flowering phenology of a population may mask responses at the individual level. For example,

Does flower visitation rates and/or reproductive success depend on the timing of flowering for the indiviual flower?

@@

![](../figures/figure_1.png){ width=100% }
**Figure 1:** Simple centroid tracking may produce erronous associations when objects move between frames. Blue shows detections in the current frame (bounding box and centroid point). Red shows centroid points for the detections in the previous frame.

### User parameters

**Max distance threshold**, **running mean**, **max disappeared**,

As the wind shifts, the flower heads changes direction. This can happen instantaneously (i.e. between two censecutive frames). As they are constrained by their stalk, there is a limit to the distance they can move.

@@

As winds shift, flowers close to the edge of the image .. may move in and out of view. If a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed. The parameter **max disappeared** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area.

@@

Setting **max disappeared** to 0 tracks objects based on the coordinates of the points in the previous frame.

The counter for number of disappeared frames is reset when a new point is associated with the track within the threshold.

@@

Mota counts tracking mismathces (shifts in a track).

@@

>>>>

---
new_session: FALSE
title: |
  **Monitoring the phenology of individual flowers using deep learning and automatic tracking**
author:
- Hjalte M. R. Mann
- Alexandros Iosifidis
- Toke T. Høye
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
fontfamily: mathpazo
fontsize: 12pt
geometry: margin = 1in
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{dcolumn}
- \usepackage{caption}
- \usepackage{float}
- \usepackage{afterpage}
- \usepackage{siunitx}
- \usepackage{amsmath}
keywords: Tracking, flowering phenology, Arctic, 
bibliography: ./library.bib
csl: ./journal-of-ecology.csl
abstract: ABSTRACT | Often simple variables will be used to describe the flowering phenology of a population of plants, e.g. onset or peak of flowering and to infer respones to climate change. Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of indiviuals. Further, we present an automatic flower tracking algorithm.
---

@@

Online tracking:
SORT
DeepSORT

@@

* **Flowering phenology**
  + Importance of studying flowering phenology
  + Responses to climate change
  + Phenology of communities, populations, individuals
  + Traditional methods for studying
  + Onset of flowering says little about true distribution
    - Even true distribition of community says little about flowering lengths of individuals and for example how it varies accross the season.
  + Difficult to study at the individual level - requires high temporal resolution and keeping track of individuals
* **Image based monitoring**
  + Automatic, high temporal resolution, remote sites
  + High temporal resolution means that we can annotate individuals and get phenology of the individual

* **Tracking**
  + Offline and online
  + Online often coupled with CNNs that attempt to distinguish individuals from each other and recognize them through frames

  + Flowers appear very similar and 
  + Many methods for offline tracking
  + Hungarian/Kahlmann filter
  ++ May not be applicable for objects that move weirdly, e.g. change directions between frames.
  + Tracking based on distance.
    - Good but has some problems
    - Two points always associated disregarding absolute distance
    - Tracks lost when objects disappear
    - Objects close to each other may swap tracks

* **Our solution**
  + Here we demonstrate a framework for automatic flower tracking and evualutation of tracking performance
  + Ground truth tracks

\newpage

For the individual plant, timing of flowering is of utmost importance. Precocious flowering means that the plant has failed to exploit the whole temporal window for accumulating resources before allocating energy to flowering. On the other hand, flowering too late limit the time for reproduction before the end of the growing season [@ELZINGA2007]. Further, flowering may need to be synchronous with pollinator activity for successful reproduction. Flowering phenology may plastically change as a response to abiotic cues in the environment, such as timing of spring, temperature, and photoperiod, but variation in flowering phenology is partly heritable and shaped by selective forces from the abiotic and biotic environment.

Monitoring of flower phenology at high temporal resolution is laboursome and time-consuming, particularly in logistically challenging environment such as the Arctic. Consequently, simple variables are often used to represent the flowering phenology of a population, such as the date for onset of flowering, often derived from weekly observations of sample plots.

Automatic image-based monitoring of flowering phenology can return phenology data for specific species at very high temporal resolution (Mann et al., in prep).

For example, phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level. For example, a shortening of flower longevity may not be directly obvious at the population level. Further, investigating the association between reproductive success and timing of flowering and flower longevity requires phenology data at the level of individuals.

Does flower visitation rates and/or reproductive success depend on the timing of flowering for the individual flower?

@@

![](../figures/figure_1.png){ width=100% }

**Figure 1:** Simple centroid tracking may produce erronous associations when objects move between frames. Blue shows detections in the current frame (bounding box and centroid point). Red shows centroid points for the detections in the previous frame.

## User parameters

As the wind shifts, the flower heads changes direction. This can happen instantaneously (i.e. between two consecutive frames). As they are constrained by their stalk, there is a limit to the distance they can move.

@@

As winds shift, flowers close to the edge of the frame may move in and out of view. If a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed. The parameter **max disappeared** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area.

@@

Setting **max disappeared** to 0 tracks objects based on the coordinates of the points in only the previous frame. The counter for number of disappeared frames is reset when a new point is associated with the track within the threshold.

@@

Mota counts tracking mismatches (shifts in a track).

@@

\newpage
# References\
